# Recommendations for Future Work {#recommendations}

Should highlight the identification of questions / outstanding guidance that were beyond our scope

- Specific questions about how well index sites within a CU represent the CU as a whole (i.e., representativeness question, but for sites within a CU instead of CUs within an SMU)

- Consideration of additional habitat-based benchmarks

- Suggestion to consider all three management levers in LRP development: harvest, hatcheries and habitat. Should LRPs be identified along those three dimensions and/or integrate information for those 3 levers?

- Standard versus hierarchical stock-recruitment models for estimating benchmarks


- We found that the projecton based LRP values were dependant on the long term exploitation rate average. Higher values of exploitation rates resulted in higher abundance LRPs. Future research could use this feature of the projection tool to identify the harvest rates that would result in aggregate LRP values that are lower than the past observed aggregate abundances. 

## Extra text on R MARKDOWN BASICS: IGNORE {#rmd-basics}

Here is a brief introduction into using _R Markdown_. _Markdown_ is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. _R Markdown_ provides the flexibility of _Markdown_ with the implementation of **R** input and output.  For more details on using _R Markdown_ see <http://rmarkdown.rstudio.com>.

Be careful with your spacing in _Markdown_ documents.  While whitespace largely is ignored, it does at times give _Markdown_ signals as to how to proceed.  As a habit, try to keep everything left aligned whenever possible, especially as you type a new paragraph.  In other words, there is no need to indent basic text in the Rmd document (in fact, it might cause your text to do funny things if you do).

## LISTS

CSAS allows ordered lists:

1. Item 1
4. Item 2

and bulleted lists:

- Item 1
- Item 2

Notice in the ordered list that I intentionally mislabeled Item 2 as number 4.  _Markdown_ automatically figures this out!  You can put any numbers in the list and it will create the list.  Check it out below.

To create sublists, just indent the values a bit (at least three spaces) and make sure there are no newlines between the items. (Here's one case where indentation is key!)

1. Item 1
1. Item 2
1. Item 3
   a. Item 3a
   a. Item 3b
      i. Item 3bi
      i. Item 3bii
      i. Item 3biii
      i. Item 3biv
      i. Item 3bv
         1. Item 3bv1
         1. Item 3bv2

## LINE BREAKS

Make sure to add white space between lines if you'd like to start a new paragraph.  Look at what happens below in the outputted document if you don't:

Here is the first sentence.  Here is another sentence.  Here is the last sentence to end the paragraph.
This should be a new paragraph.

*Now for the correct way:*

Here is the first sentence.  Here is another sentence.  Here is the last sentence to end the paragraph.

This should be a new paragraph.

## R CHUNKS

When you click the **Knit** button above a document will be generated that includes both content as well as the output of any embedded **R** code chunks within the document. You can embed an **R** code chunk like this (`cars` is a built-in **R** dataset):

```{r cars2}
summary(cars)
```
## INLINE CODE

If you'd like to put the results of your analysis directly into your discussion, add inline code like this:

> The `cos` of $2 \pi$ is `r cos(2*pi)`.

Another example would be the direct calculation of the standard deviation:

> The standard deviation of `speed` in `cars` is `r sd(cars$speed)`.

One last neat feature is the use of the `ifelse` conditional statement which can be used to output text depending on the result of an **R** calculation:

> `r ifelse(sd(cars$speed) < 6, "The standard deviation is less than 6.", "The standard deviation is equal to or greater than 6.")`

Note the use of `>` here, which signifies a quotation environment that will be indented.

As you see with `$2 \pi$` above, mathematics can be added by surrounding the mathematical text with dollar signs.  More examples of this are in [Mathematics and Science] if you uncomment the code in [Math].

## INCLUDING PLOTS

You can also embed plots.  For example, here is a way to use the base **R** graphics package to produce a plot using the built-in `pressure` dataset:
```{r pressure2, echo=FALSE, cache=TRUE}
plot(pressure)
```
Note that the `echo=FALSE` parameter was added to the code chunk to prevent printing of the **R** code that generated the plot.  There are plenty of other ways to add chunk options.  More information is available at <http://yihui.name/knitr/options/>.

Another useful chunk option is the setting of `cache=TRUE` as you see here.  If document rendering becomes time consuming due to long computations or plots that are expensive to generate you can use knitr caching to improve performance.  Later in this file, you'll see a way to reference plots created in **R** or external figures.

## LOADING AND EXPLORING DATA

Included in this template is a file called `flights.csv`.  This file includes a subset of the larger dataset of information about all flights that departed from Seattle and Portland in 2014.  More information about this dataset and its **R** package is available at <http://github.com/ismayc/pnwflights14>.  This subset includes only Portland flights and only rows that were complete with no missing values.  Merges were also done with the `airports` and `airlines` data sets in the `pnwflights14` package to get more descriptive airport and airline names.

\  

We can load in this data set using the following command (`read_csv()` is from the package `readr`):
```{r load_data2, echo=TRUE}
flights <- read_csv(file.path("data", "flights.csv"))
```  

\  

*Before we continue, note that to insert extra blank lines in text as we did above and below, we need to have a line with a backslash followed by two spaces. Look carefully at those lines in the code editor.*

\  

All your tabular data should be in `.csv` files and placed in the `data` directory. The function `here::here()` should always be used when loading data into the project because it will prepend the full root directory of the project onto the directories and files inside the call. This gives consistent file structure access across multiple machines and platforms.

The data is now stored in the data frame called `flights` in **R**.  To get a better feel for the variables included in this dataset we can use a variety of functions.  Here we can see the dimensions (rows by columns) and also the names of the columns.

```{r str2}
dim(flights)
names(flights)
```
\  

Another good idea is to take a look at the dataset in table form.  With this dataset having more than 50,000 rows, we won't explicitly show the results of the command here.  I recommend you enter the command into the Console **_after_** you have run the **R** chunks above to load the data into **R**. To do that, you can use the command `View(flights)`.

A simple and effective way to see the entire structure in one call is to use the `glimpse()` command like this:
```{r glimpse_flights2, echo = TRUE}
glimpse(flights)
```

```{r view_flights2, eval=FALSE}
View(flights)
```

While not required, it is highly recommended you use the `dplyr` package to manipulate and summarize your data set as needed.  It uses a syntax that is easy to understand using chaining operations.  Below I've created a few examples of using `dplyr` to get information about the Portland flights in 2014.  You will also see the use of the `ggplot2` package, which produces beautiful, high-quality academic visuals.

We begin by checking to ensure that needed packages are installed and then we load them into our current working environment:

```{r load_pkgs2, message=FALSE, echo=TRUE}
# List of packages required for this analysis
pkg <- c("dplyr", "ggplot2", "knitr", "bookdown", "devtools")
# Check if packages are not installed and assign the
# names of the packages not installed to the variable new.pkg
new.pkg <- pkg[!(pkg %in% installed.packages())]
# If there are any packages in the list that aren't installed,
# install them
if (length(new.pkg)) {
  install.packages(new.pkg, repos = "http://cran.rstudio.com")
}
library(csasdown)
```

\clearpage

The example we show here does the following:

- Selects only the `carrier_name` and `arr_delay` from the `flights` dataset and then assigns this subset to a new variable called `flights2`.

- Using `flights2`, we determine the largest arrival delay for each of the carriers.

```{r max_delays2}
library(dplyr)
flights2 <- flights %>%
  select(carrier_name, arr_delay)
max_delays <- flights2 %>%
  group_by(carrier_name) %>%
  summarize(max_arr_delay = max(arr_delay, na.rm = TRUE))
```

A useful function in the `knitr` package for making nice tables in _R Markdown_ is called `kable`.  It is much easier to use than manually entering values into a table by copying and pasting values into Excel or LaTeX.  This again goes to show how nice reproducible documents can be! (Note the use of `results="asis"`, which will produce the table instead of the code to create the table.)  The `caption.short` argument is used to include a shorter title to appear in the List of Tables.

Within CSAS you should use the function `csasdown::csas_table()` instead. It just calls the function `kable()`, but it adds some default argument values that will make the tables render correctly in both LaTeX and in Word documents.


```{r maxdelays2, results="asis"}
library(knitr)
csas_table(max_delays,
  col.names = c("Airline", "Max Arrival Delay"),
  caption = "Maximum Delays by Airline"
)
```

The last two options make the table a little easier-to-read.

Let's make a plot of the data:


```{r march3plot2, fig.height=3, fig.width=6, fig.cap="My caption."}
library(ggplot2)
flights %>%
  ggplot(aes(x = dep_time, y = arr_delay)) +
  geom_point()
```

\ 
Also, making equations in text:


\begin{equation}
  1 + 1 = 2
  (\#eq:reaction)
\end{equation}

\clearpage 
By far the easiest way to present tables in your report is to store the contents of the table in a CSV or Excel file, then read that file in to your R Markdown document as a data frame. Then you can style the table with the `kable` function, or functions in the [kableExtra](https://cran.r-project.org/web/packages/kableExtra/index.html) pacakge. We recommend using `csasdown::csas_table()`, which simply passes some necessary arguments to `knitr::kable()` so that the tables will render correctly for CSAS.

In addition to the tables that can be automatically generated from a data frame in R that you saw in [R Markdown Basics] using the `kable` function, you can also create tables using pandoc. (More information is available at <http://pandoc.org/README.html#tables>.)  This might be useful if you don't have values specifically stored in R, but you'd like to display them in table form.  Below is an example.  Pay careful attention to the alignment in the table and hyphens to create the rows and columns. Generally we don't recommend this approach of typing the table directly into your R Markdown document.

----------------------------------------------------------------------------------
  Factors                    Correlation between Parents & Child      Inherited
------------------------- ----------------------------------------- --------------
  Education                                -0.49                         Yes
  
  Socio-Economic Status                     0.28                        Slight   
  
  Income                                    0.08                          No
  
  Family Size                               0.18                        Slight
  
  Occupational Prestige                     0.21                        Slight
------------------------- ----------------------------------------- --------------
Table: (\#tab:inher) Correlation of Inheritance Factors for Parents and Child 

We can also create a link to the table by doing the following: Table \@ref(tab:inher).  If you go back to [Loading and exploring data] and look at the `kable` table, we can create a reference to this max delays table too: Table \@ref(tab:maxdelays). The addition of the `(\#tab:inher)` option to the end of the table caption allows us to then make a reference to Table `\@ref(tab:label)`. Note that this reference could appear anywhere throughout the document after the table has appeared.  

\clearpage

<!-- clearpage ends the page, and also dumps out all floats.
  Floats are things like tables and figures. -->

\begin{landscapepage}
```{r widelong, results = "asis"}
yrs <- 1:45
ages <- 1:15
widelong <- matrix(round(rnorm(length(yrs) * length(ages), 20), 4),
                        nr = length(yrs),
                        nc = length(ages),
                   dimnames = list(year = yrs, age = ages))
widelong_df <- data.frame(cbind(yrs, widelong))

# In landscape tables, you can change the font size if the table
# is too large and overlaps the header and footer lines.
csasdown::csas_table(widelong_df,
                     col.names = c("Year", as.character(ages)),
                     caption = "A long and wide table",
                     format = "latex",
                     font_size = 8)
```
\end{landscapepage}
\clearpage

## FIGURES

If your report has a lot of figures, _R Markdown_ might behave better for you than that other word processor. One perk is that it will automatically number the figures accordingly in each chapter. You'll also be able to create a label for each figure, add a caption, and then reference the figure in a way similar to what we saw with tables earlier.  If you label your figures, you can move the figures around and _R Markdown_ will automatically adjust the numbering for you.  No need for you to remember!  So that you don't have to get too far into LaTeX to do this, a couple R functions have been created for you to assist.  You'll see their use below.

<!--
One thing that may be annoying is the way _R Markdown_ handles "floats" like tables and figures (it's really \LaTeX's fault). \LaTeX\ will try to find the best place to put your object based on the text around it and until you're really, truly done writing you should just leave it where it lies. There are some optional arguments specified in the options parameter of the `label` function.  If you need to shift your figure around, it might be good to look here on tweaking the options argument:  <https://en.wikibooks.org/wiki/LaTeX/Floats,_Figures_and_Captions>

If you need a graphic or tabular material to be part of the text, you can just put it inline. If you need it to appear in the list of figures or tables, it should be placed in a code chunk.
-->

In the R chunk below, we will load in a picture stored as `uw.png` in our main directory.  We then give it the caption of "UW logo", the label of "uwlogo", and specify that this is a figure.  Make note of the different R chunk options that are given in the R Markdown file (not shown in the knitted document).

```{r example-fig, fig.cap="My caption"}
plot(1:10)
```

Here is a reference to the figure: Figure \@ref(fig:example-fig).  Note the use of the `fig:` code here.  By naming the R chunk that contains the figure, we can then reference that figure later as done in the first sentence here.  We can also specify the caption for the figure via the R chunk option `fig.cap`.

\clearpage 

<!-- starts a new page and stops trying to place floats such as tables and figures -->

Below we will investigate how to save the output of an R plot and label it in a way similar to that done above.  Recall the `flights` dataset from Section \@ref(rmd-basics).  (Note that we've shown a different way to reference a section here.)  We will next explore a bar graph with the mean flight departure delays by airline from Portland for 2014.  Note also the use of the `scale` parameter which is discussed on the next page.

```{r delaysboxplot, warnings=FALSE, messages=FALSE, fig.cap="Mean Delays by Airline", fig.width=6}
flights %>% group_by(carrier) %>%
  summarize(mean_dep_delay = mean(dep_delay)) %>%
  ggplot(aes(x = carrier, y = mean_dep_delay)) +
  geom_bar(position = "identity", stat = "identity", fill = "red")
```

Here is a reference to this image: Figure \@ref(fig:delaysboxplot).

A table linking these carrier codes to airline names is available at <https://github.com/ismayc/pnwflights14/blob/master/data/airlines.csv>.

\clearpage

## FOOTNOTES AND ENDNOTES

You might want to footnote something. ^[footnote text] The footnote will be in a smaller font and placed appropriately. Endnotes work in much the same way. 

## BIBLIOGRAPHIES

Of course you will need to cite things, and you will probably accumulate an armful of sources. There are a variety of tools available for creating a bibliography database (stored with the .bib extension).  In addition to BibTeX suggested below, you may want to consider using the free and easy-to-use tool called Zotero. Some Zotero documentation is at <http://libguides.reed.edu/citation/zotero>.  In addition, a tutorial is available from Middlebury College at <http://sites.middlebury.edu/zoteromiddlebury/>.

R Markdown uses pandoc (<http://pandoc.org/>) to build its bibliographies.  One nice caveat of this is that you won't have to do a second compile to load in references as standard LaTeX requires. To cite references in your report (after creating your bibliography database), place the reference name inside square brackets and precede it by the "at" symbol.  For example, here's a reference to a book about worrying:  [@francis2011].  This `francis2011` entry appears in a file called `refs.bib` in the `bib` folder.  This bibliography database file was created by a program called BibTeX.  You can call this file something else if you like (look at the YAML header in the main .Rmd file) and, by default, is to placed in the `bib` folder. Another example: @edwards2013 shows that... And another example: [e.g., @francis2011; @edwards2013] ...

For more information about BibTeX and bibliographies, see (<http://web.reed.edu/cis/help/latex/index.html>). There are three pages on this topic:  _bibtex_ (which talks about using BibTeX, at <http://web.reed.edu/cis/help/latex/bibtex.html>), _bibtexstyles_ (about how to find and use the bibliography style that best suits your needs, at <http://web.reed.edu/cis/help/latex/bibtexstyles.html>) and _bibman_ (which covers how to make and maintain a bibliography by hand, without BibTeX, at <http://web.reed.edu/cis/help/latex/bibman.html>). The last page will not be useful unless you have only a few sources.

**Tips for Bibliographies**

- The cite key (a citation's label) needs to be unique from the other entries.
- When you have more than one author or editor, you need to separate each author's name by the word "and" e.g. `Author = {Noble, Sam and Youngberg, Jessica},`.
- Bibliographies made using BibTeX (whether manually or using a manager) accept LaTeX markup, so you can italicize and add symbols as necessary.
- To force capitalization in an article title or where all lowercase is generally used, bracket the capitalized word in curly braces. E.g. `{Pacific Cod}`.

CSAS style requires the bibliography to be before the appendices and before tables and figures if you keep those together at the end of your document.

See the next file `04-references.Rmd` which will make the bibliography appear wherever that file is listed in `_bookdown.yml`.

## ADDITIONAL RESOURCES

- _Markdown_ Cheatsheet - <https://github.com/adam-p/markdown-here/wiki/Markdown-Cheatsheet>

- _R Markdown_ Reference Guide - <https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf>

- Introduction to `dplyr` - <https://cran.rstudio.com/web/packages/dplyr/vignettes/introduction.html>

- `ggplot2` Documentation - <http://docs.ggplot2.org/current/>
